---
title: "Fixed Income:Black_Derman_Toy Model"
author: "YiTao Hu, Charles Rambo, Junyu(Kevin) Wu, Jin (Jane) Huangfu"
date: "20/02/2020"
output: pdf_document
---

```{r}
#import data
library(readxl)
vol_ts=read_excel("Homework 6 voldat.xlsx",col_names = FALSE)
vol_ts=vol_ts$...1
True_DTs = read_excel("Homework 6 pfilea.xlsx", col_names = FALSE)
True_DTs=True_DTs$...1
```
### Question 1
##initizliation of interest rate tree

We first need to randomly initialize the short rate $r_t^*$ at each point of time t. 
```{r}
#initizlize R*
steps=29
r_stars=abs(rnorm(n = (steps+1))/100)
```

From $r_t^*$, we can map all the interest rate tree with the folloing formula:
$$r_t^{id}=r_t^*e^{-2i\sigma_t\sqrt{0.5}}$$
where i is the number of doward movements from the top state. We can define a function to do the computation. 

```{r}

#define a function to compute R^id from R^*
get_Rid=function(r_star,id,vol){
  rid=r_star*exp(-2*id*vol*0.5^0.5)
  return (rid)
}
get_all_Rids=function(r_stars,vol_ts){
  #initialize the rate tree dataframe 
  rateTree=data.frame(matrix(0,nrow = length(r_stars),ncol = length(r_stars)))
  colnames(rateTree)=seq(0,(length(r_stars)-1)/2,0.5)
  rateTree[1,]=r_stars
  
  #compute and store the value of rateTree
for (t in 2:ncol(rateTree)){
  for(i in 2:t){
    rateTree[i,t]=get_Rid(r_star = r_stars[t],id = i,vol = vol_ts[(t-1)])
  }
}
  return (rateTree)
}

Rates_tree=get_all_Rids(r_stars = r_stars,vol_ts = vol_ts)
```

## Mapping interest rate tree to discount factor tree and compute recursively discount factor curve

Given the short interest rate tree, we can map the interest rate tree to a discount factor tree of 0.5 year maturity at each state $s$ and each time step $t$ because short term discount factor is an one-to-one mapping relationship. In particular, we will use the following formula to compute discount factor
$$D^{id}(t,0.5)=\frac{1}{1+r^{id}/2}$$
where $id$ refers to the number of downward movements or state, $t$ refers to the time step.
```{r}
get_short_DT_tree=function(Rates_tree){
  DT_short_tree=1/(1+Rates_tree/2)
  return (DT_short_tree)
}
DT_short_tree=get_short_DT_tree(Rates_tree = Rates_tree)
```

Given the short-term discount factor tree, we can compute the cash flow tree or discount factor tree with various maturity using the following backward recursive scheme for each $D(0,T)$ of T=0.5,1,...,15
$$D^{s}(t,T)=D^s(t,0.5)\frac{D^{su}(t+0.5,T-0.5)+D^{sd}(t+0.5,T-0.5)}{2}$$
where $s$ denotes the state, $su$ denotes upward state and $sd$ denotes doward state one step later. This formula, in essence, is equivalent to the recusive scheme in the TA notes with interest rate. 
```{r}
#define a function to compute each individual D(0,T)
compt_D0T=function(DT_short_tree,Mar){
  #initialize the matrix of D^s(t,T) matrix 
  DsT_mat=DT_short_tree[1:Mar,1:Mar]
  #perform the backward inductive scheme
  for (t in Mar:2){
    for (i in 1:(t-1)){
      DsT_mat[i,(t-1)]=DsT_mat[i,(t-1)]*mean(c(DsT_mat[i,t],DsT_mat[(i+1),t]))
    }
  }
  return (DsT_mat[1,1])}
#define a function to compute all the D(0,T)
comp_all_D0T=function(DT_short_tree){
  #initialize a vector of 30 discount factor at time 0
  fitted_DTs=rep(0,ncol(DT_short_tree))
  fitted_DTs[1]=DT_short_tree[1,1]
  for (Mar in 2:length(fitted_DTs)){
    fitted_DTs[Mar]=compt_D0T(DT_short_tree = DT_short_tree,Mar = Mar)
  }
return (fitted_DTs)
}

fitted_DTs=comp_all_D0T(DT_short_tree = DT_short_tree)
```

## Define Cost function and Perform Optimization to get the numerical solution
Because it is complicated to solve all the $r_t^*$ analytically, we turn to find its numerical solution by minimizing the Mean Average Error between the fitted DTs and the true DTs. 

We, first, need to defined a weighted sum error function (because the algorithm overweighs the first parameter and underweighs the last one, we want to do a sigmoid-like transformation to reverse this weighting):
```{r}
compt_weightedError=function(True_values,fitted_values){ 
  WE=sum((1/(1+exp(-seq(-1.5,1.5,3.0/29))))*abs(fitted_values-True_values))
  #WE=sum(seq(2,3,1/29)*abs(fitted_values-True_values))
  return (WE)
}
WE=compt_weightedError(True_values = True_DTs,fitted_values = fitted_DTs)
```

Then, we can assimble all the function into a pipeline as a cost function or optimization objective. 
```{r}
Cost_Fun=function(r_stars,vol_ts,True_DTs){
  Rates_tree=get_all_Rids(r_stars = r_stars,vol_ts = vol_ts)
  DT_short_tree=get_short_DT_tree(Rates_tree = Rates_tree)
  fitted_DTs=comp_all_D0T(DT_short_tree = DT_short_tree)
  WE=compt_weightedError(True_values = True_DTs,fitted_values = fitted_DTs)
  return (WE)
}
Cost_Fun(r_stars = r_stars,vol_ts = vol_ts ,True_DTs = True_DTs)
```
Finally, we can feed the Cost Function into optimizer and try to find the optimal numerical solution. 
```{r}
#Num_sol=optim(par = r_stars,fn = Cost_Fun,method = "BFGS",vol_ts=vol_ts,True_DTs=True_DTs,control = list(maxit=100),lower = rep(0,30))
Num_sol=nlminb(start = r_stars,
               objective =Cost_Fun,vol_ts=vol_ts,True_DTs=True_DTs,lower = rep(0,30),control = list(trace=1,iter.max=600))
R_star_num=Num_sol$par
R_star_num
```

Once we got our numerical solution for $r_t^*$, we can again compute the whole tree.
```{r}
Opt_tree=get_all_Rids(r_stars = R_star_num,vol_ts = vol_ts)
```



### Question 2

Compute the expected value of the short rate tree at each time step, and plot the risk_neural short rate curve.Because the rates follow binary distribution.
The martingale probablity of each rate should be 
$$C^{iu}_n(\frac{1}{2})^n$$
where n is the number of steps, i is the number of upward movement.

```{r}
#turn all zero values into NA
Opt_tree[Opt_tree==0]=NA
#compute a martinagle probability matrix 
Q_prob_mat=Opt_tree
#compute the martingale probability of each node
for (t in 1:ncol(Q_prob_mat)){
  for (i in 1:t){
    Q_prob_mat[i,t]=choose(n = t-1,k = i-1)*(1/2)^(t-1)
  }
}
#compute the martingale probability 
Ex_R_q=colSums(Opt_tree*Q_prob_mat,na.rm = T)
#plot the risk_neural short rate curve
plot(y=Ex_R_q,x=seq(0,14.5,0.5),type='l',main = 'Risk Neural Expected short rate Curve',
     ylab = 'Rate',xlab = 'Maturity')
```

Recall the formula of forward rate in lecture1 from discount factor, we can compute the short-term (0.5 year) forward rates at each time step using the following formula:
$$2tf_{0.5}=2\frac{(1+r_{t+0.5}/2)^{t+0.5}}{(1+r_t/2)^t}-1=2\frac{D(0,t)}{D(0,t+0.5)}-1$$
```{r}
#compute the forward rates and plot it over the martingale expected rate
forward_rates=Ex_R_q
forward_rates[1]=2*(1/True_DTs[1]-1)
forward_rates[2:30]=2*(True_DTs[-length(True_DTs)]/True_DTs[-1]-1)
#plot the short-term rate overlay on the risk-neutral expected short rate
plot(Ex_R_q,type='l',main = 'Martingale expected rate over Forward rate Curve',
     ylab = 'Rate',xlab = 'Maturity',col='red')
lines(x = forward_rates,col='blue')
legend('topleft', legend = c("Q Expected Rates", "Forward Rates"),
col = c("red", "blue"), lty = 1)
```

