---
title: "Fixed_Income_stringModel_YiTaoHu"
author: "YiTao Hu"
date: "26/02/2020"
output:
  pdf_document: default
  html_document: default
---
```{r message=FALSE, warning=FALSE}
#first we import data and some useful libraries
library(derivmkts)
library(LSMonteCarlo)
library(readxl)
library(readr)
D_0_T=read_excel("Homework 7 pfilea.xlsx", col_names = FALSE)
D_0_T=D_0_T$...1
sigmas=read_excel("Homework 7 sigma.xlsx",      col_names = FALSE)
sigmas=sigmas$...1
Corr_mat=read_csv("Homework 7 corrin.csv", 
    col_names = FALSE)
U=read_csv("Homework 7 corchol.csv", 
     col_names = FALSE)
U=as.matrix(U)
```
##Q1
#Solve for r* and compute recusive scheme for dD(t,T)
Recall dynamics of the bond price under string model
$$dD(t,T)=r_t^*D(t,T)dt+\sigma(T-t)D(t,T)dW(t,T)$$

where $W(t,T)$ are correlated Brownian Motion at time t. in other words, $d\vec{W_t}$ is a multivariate normal variavle $MVN(0,\Sigma)$
where $\Sigma$ is the var-cov matrix of $MVN$

First,we define a function to compute $d\vec{W_t}$. 
$$\vec{dW_t}=\vec{W_{t+dt}}-\vec{W_t}=\sqrt{dt}U^T\vec{Z}$$

where $U^T$ is the Cholesky root of Correlation matrix and $\vec{Z}$ is a standard MVN
```{r}
compute_dWt=function(cho_corr,dt){
  dWt=sqrt(dt)*cho_corr%*%rnorm(nrow(as.matrix(cho_corr)))
  return(dWt)
}
dWt=compute_dWt(cho_corr = U[-20,-20],dt = 0.5)
```

Then we need to define a function to compute r*
```{r}
compt_rf=function(D_0.5){
  rf=2*(1/D_0.5-1)
  return(rf)
}
rf=compt_rf(D_0.5 = D_0_T[1])
```

From the formula above, we can get our extrapolation scheme:
$$D(t+dt,T)=D(t,T)+dD(t,T)=D(t,T)+r_t^*D(t,T)dt+\sigma(T-t)D(t,T)dW(t,T)$$
```{r}
#define a function to compute D(t+dt,T)
comput_Dtforward=function(DT,D_0.5,sigmas,dt,cho_corr){
  rf=compt_rf(D_0.5 = D_0.5)
  dWt=compute_dWt(cho_corr = cho_corr,dt = dt )
  D_tplus=DT+rf*DT*dt+sigmas*DT*dWt
  return(D_tplus)
}

```
#Compute one D(t,T) trajectory
Now we can compute one particular trajectories of the whole DT function series using the recusive scheme defined above. 
```{r}
#define a function to compute one particular senerio
cho_corr=U
senoro_sim=function(D0T,sigmas,cho_corr){
  #initialize the scenario matrix 
  scenario_df=data.frame(matrix(0,nrow = 20,ncol = 20))
  colnames(scenario_df)=seq(0,9.5,0.5)
  scenario_df[,1]=D0T
  #simulate the trajectories
  for (t in 2:20){
    scenario_df[t:20,t]=comput_Dtforward(DT = scenario_df[t:20,(t-1)],D_0.5 = scenario_df[(t-1),(t-1)],sigmas = sigmas[1:(21-t)],dt = 0.5,cho_corr = cho_corr[1:(21-t),1:(21-t)])
  }
  return(as.matrix(scenario_df))
}
```
Then, we can perform 10,000 scenarios simulation
```{r}
batch_sim=function(D0T,sigmas,scenario_num=10000)
  {
  #initialize the simulation
  sim_array=array(data = 0,dim = c(20,20,scenario_num))
  #because paralell computing exceeds the CPU limit, here we use for loop to do simulation
  for(i in 1:scenario_num){
    sim_array[,,i]=senoro_sim(D0T = D0T,sigmas = sigmas,cho_corr = cho_corr)
  }
  return(sim_array)
}
sim_out=batch_sim(D0T=D_0_T[1:20],sigmas=sigmas[1:20],scenario_num=10000)
#Then,we can look at the structure's average time decay across different scenerios
Ave_sim_decay=apply(sim_out, MARGIN =c(1,2), FUN = mean)
Ave_sim_decay
```
##2
Here, we need to compute the forward par rates five years ahead from the initial term structure.
$$C_{forward}=2[\frac{100D(N)-100D(N+M)}{\sum^{2M}_{i=1}D(N+i/2)}]$$

```{r}
Forward_DTs=D_0_T[11:21]
Forward_par=2*(100*Forward_DTs[1]-100*Forward_DTs[2:11])/cumsum(Forward_DTs[2:11])
Forward_par
```


##3
We first need to define a function to map discount factor to par rates.
```{r}

#define a function to get the Discount factor decay for T=0.5:5
get_DT_decay=function(sim_mat)
{
  DTdecay=matrix(data = 0,nrow = 11,ncol = 10)
  Trim_mat=sim_mat
  for (T in 1:10){
    DTdecay[,T]=diag(Trim_mat)[1:11]
    Trim_mat=Trim_mat[-1,-ncol(Trim_mat)]
  }
  DTdecay=t(DTdecay)
  rownames(DTdecay)=seq(0.5,5,0.5)
  colnames(DTdecay)=seq(0,5,0.5)
  return(DTdecay)
}
```
Compute the par rates at each time step
```{r}
compt_par=function(DT){
  ParRates=2*(100-100*DT)/cumsum(DT)/100
  return(ParRates)
}
#assible the two function above into one 
get_par_rates=function(sim_mat){
  DTs=get_DT_decay(sim_mat)
  Par_decay=apply(DTs, 2, compt_par)
  return(Par_decay)
}
#compute the terminal payoff at each scenarios 
get_batch_par=function(sim_asset){
  #initialize the par simulation arrary
  par_array=array(data = 0,dim = c(10,11,dim(sim_asset)[3]))
  #because paralell computing exceeds the CPU limit, here we use for loop to do simulation
  for(i in 1:dim(sim_asset)[3]){
    par_array[,,i]=get_par_rates(sim_asset[,,i])
  }
  return(par_array)
}
#compute all par rates decay based on all simulation
par_rates_batch=get_batch_par(sim_asset = sim_out)
```


Recall the future price of the contract is the risk-neutral expected value of the underlying asset:
$$K=E^Q(B_T)$$
where $B_T$ in this case is the price of the bond to deliver at terminal time T.

First,we need to mapping par rates dynamics to bond price at time T.
```{r}
#compute bond price at terminal time T=5 for each scenerio and each maturity
bond_value=data.frame(matrix(0,nrow = 10,000,ncol = 10))
colnames(bond_value)=seq(0.5,5,0.5)
for (i in 1:dim(par_rates_batch)[3]){
    for (mar in 1:10){
      bond_value[i,mar]=bondpv(coupon = Forward_par[mar],mat = (mar/2),yield =  par_rates_batch[mar,11,i],principal = 100,freq = 2)
    }
}
#select out the cheapest bond for each scenerio
B_T_min=apply(bond_value[,seq(2,10,2)], 1, min)
#compute the risk-neutral expected value 
K=mean(apply(bond_value[,seq(2,10,2)], 2, mean))
K
```

Recall the price of any rate-related derivative should be the risk-neutral expected value of discounted final cash flow.
$$Option=E^Q(e^{-\int^T_0Rsd_s}(K-B_{Tcheap}))$$
where both the bond to deliver $B_{Tcheap}$and the short rates $Rs$ are stochastic.
```{r}
#grab the paths of short rates for all timestep
Discount_sim=apply(sim_out, 3, diag)[1:10,]
#multiple the short rate to get total discount factor for each scenoro
Disount_T=apply(Discount_sim, 2, cumprod)[10,]
#compute the option price 
Opt_val=mean(Disount_T*(K-B_T_min))
Opt_val
```

