---
title: "Fixed Income HW4"
author: "YiTao Hu, Charles Rambo, Junyu(Kevin) Wu, Jin (Jane) Huangfu"
date: "07/02/2020"
output: pdf_document
---

```{r}
#import the data
Rate_paths=data.frame(matrix(data = c(0.04,0.04,0.04,0.04,0.04,0.05,0.03,0.04,0.07,0.02,0.07,0.02,0.05,0.06,0.08,0.06,0.04,0.07,0.04,0.09,0.03,0.07,0.08,0.04,0.12,0.06,0.06,0.10,0.05,0.09),nrow = 5,ncol = 6))
colnames(Rate_paths)=seq(0,5,1)
Rate_paths
```
#1 
Recall from the slides, under risk neutral valuation, the PV formula:
$$PV=E^Q_0[e^{-AT}\Phi(X_T)]$$
where A is the average rate from 0 to T and $\Phi(X_T)$ is the payoff function at T
```{r}
valuate_risk_neutral=function(paths,FUN=100,...){
   #initialization
    DerivPrice=rep(0,ncol(paths)-1)
    for (mar in 2:ncol(paths)){
    trajectories=paths[,1:mar]
    A=rowMeans(trajectories)
    DerivPrice[mar-1]=mean(exp(-A*(mar-1))*FUN(trajectories,...))
  }
  DerivPrice=data.frame(t(DerivPrice))
  colnames(DerivPrice)=seq(1,(ncol(paths)-1),1)
  return(DerivPrice)
}
```

Put into code, the prices of ZCB would be
```{r}
#wrirte the payoff function of ZCBs
ZCP_Phi=function(df)
{
  return(100)
}
valuate_risk_neutral(paths = Rate_paths,FUN = ZCP_Phi)
```



#2
The payoff function of a in interest rate cap would be:
$$Max(0,R_T-0.05)$$
Write the payoff function of interest rate cap
```{r}
Phi_cap=function(df,K){
  Payoff=pmax(0,df[,ncol(df)]-K)
  return(Payoff)
}
```

Compute the interest rate cap:
```{r}
valuate_risk_neutral(paths = Rate_paths,FUN = Phi_cap,K=0.05)
```

#3
The payoff function of a in interest rate floor would be:
$$Max(0,K-R_T)$$
Write the payoff function:
```{r}
Phi_floor=function(df,K){
  Payoff=pmax(0,K-df[,ncol(df)])
  return(Payoff)
}
```

Compute the interest rate cap:
```{r}
valuate_risk_neutral(paths = Rate_paths,FUN = Phi_floor,K=0.07)
```

#4
Compute the price of a five year call on short interest rate:
```{r}
valuate_risk_neutral(paths = Rate_paths,FUN = Phi_cap,K=0.065)[ncol(Rate_paths)-1]
```
Compute the price of a five year put on short interest rate:
```{r}
valuate_risk_neutral(paths = Rate_paths,FUN = Phi_floor,K=0.065)[ncol(Rate_paths)-1]
```
Therefore a a five year put on short term rate is more valuable. 


#5
The payoff of the Asian option should be:
$$Max(0,\bar{R}-K)$$ 
for call and 
$$Max(0,K-\bar{R})$$ 
for put. 

Write the payoff fuction
```{r}
Phi_Asian_cap=function(df,K){
  Payoff=pmax(0,rowMeans(df)-K)
  return(Payoff)
}
Phi_Asian_floor=function(df,K){
  Payoff=pmax(0,K-rowMeans(df))
  return(Payoff)
}
```

Compute the price of the Asian call:
```{r}
valuate_risk_neutral(paths = Rate_paths,FUN = Phi_cap,K=0.06)[ncol(Rate_paths)-1]
```
Compute the price of the Asian put:
```{r}
valuate_risk_neutral(paths = Rate_paths,FUN = Phi_Asian_floor,K=0.06)[ncol(Rate_paths)-1]
```
The asian put is more valuable in this case

#6
The standard deviation of the short-term rate at year 5:
```{r}
sd(Rate_paths[,6])
```
The standard deviation of the average short-term rate year 5
```{r}
sd(rowMeans(Rate_paths))
```

Obviously, the volatility of average short-term rate is much smaller than that of terminal short-term rate. Because Option price is positively correlated the volatility of underlying asset, the European call is more valuaable than Asian Call.

#7
In the lecture, we proved that, by the condition of no abitrage, the future prices at time 0 must equal to the martingale expectation of the underlying asset at maturity T.
$$K=E^Q_0(S_T)$$
Therefore, the futures prices for short-term rate (Eurodollar futures) should be:
```{r}
colMeans(Rate_paths)
```






